---
title: "Homework 5"
author: "Natasha Hannan"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rpart)
library(rpart.plot)

mydata=read.csv("~/BI612-homework/crimedata.csv")
mydata2=read.csv("~/BI612-homework/crimedata10.csv")
```

```{r}
summary(mydata)
```

```{r}
set.seed(1337)
train=sample(x = nrow(mydata), size = 47, replace = FALSE)
train.data=mydata[sample(x = nrow(mydata), size = 47, replace = FALSE),] 
```

#3 Regression Tree Summary
```{r}
m.rpart=rpart(CrimeRate~Youth+Education+ExpenditureYear+Males+StateSize+YouthUnemployment+BelowWage, data = train.data) 
summary(m.rpart)
```
In this regression tree the most important variables were Expenditure Year, Youth and Below Wage.

#4
```{r}
rpart.plot(m.rpart, digits = 3, fallen.leaves = TRUE)
```
This diagram is showing that when it comes to crime rate 48.9% of the population has an expenditure year that is smaller than 77 and 51.1% that is greater than 77. Of the population  that has an expenditure year smaller than 77, 25.5% have a state size smaller than 23 while 2.4% have a state size greater than 23. Of the population that has an expenditure year greater than 77, 23.4% has less then 131 young males while 27.7% has more than 131 young males. This shows the likelyhood of crime depending on expenditure year, state size and youth.

#5
mean crime rate per group:
ExpenditureYear < 77 & StateSize < 23 = 72.5
ExpenditureYear < 77 & StateSize > 23 = 97.6
ExpenditureYear > 77 & Youth < 131 = 105
ExpenditureYear > 77 & Youth > 131 = 133

#6
Yes, some predictor variables were excluded from the model. I believe that this is due their lack of importance on the outcome. 


#7
```{r}
p.rpart=predict(m.rpart, mydata2)
p.rpart
```

#8
```{r}
cor(p.rpart, mydata2,method="pearson")
```
The correlation coefficient from this analysis is 0.6175

#9 
```{r}
MAE <- function(actual, predicted)  {
  mean(abs(actual - predicted))
}

MAE(predicted = p.rpart,actual = mydata2$CrimeRate)
```
I do not believe that this model is good at predicting because it is indicating that our model is predicting on average 24.329 away from the actual value. 

#10
```{r}
crime.test=mydata$CrimeRate
actual=crime.test
MAE2 <- function(data,indices)  {
  d<-data[indices]
  return(mean(abs(actual - d)))
}
library(boot)
guesses=boot(data=crime.test, statistic=MAE2, R=1000)
{hist(guesses$t)
abline(v=mean(guesses$t),col="red")}
mean(guesses$t)
p.value=length(which((guesses$t<0.5198)==T))/1000
p.value
```

#11 the mean absolute error when crime rates are randomly assigned is 32.57256 which is greater than the error from the model. 

#12 There is no significant different from the mean absolute error of the model and random become the p-value was 0. 

